Solving Combinatorial Puzzles with Parallel Evolutionary Algorithms \thanks{This work was supported by a private funding of Velbazhd Software LLC.}

### 2019-02-20 Solving-Combinatorial-Puzzles-with-Parallel-Evolutionary-Algorithms/paper.tex
Some corrections were done by Rumen.

A parallel implementation of a genetic algorithm based solver of the Rubik's cube was implemented by Balabanov in \cite{balabanov01} and it was presented in \cite{balabanov02}. Rubik's cube was invented and introduced by Erno Rubik in the 70s of the 20th century. After its creation the cube became the most popular combinatorial puzzle all over the world. In its original version it has 3x3x3 cubical segments. There are stickers in six different colors on each subcube square of the exposed sides. Each of the six planes (3x3x1) can be rotated in 90, 180, 270 or 360 degrees, relative to the other part of the puzzle. In the original initial state all sides of the cube are in single color. Scumbling of the puzzle is done by many random rotations of the (3x3x1) sides. The optimization task is to restore the cube in its original state. Such combinatorial optimization problem is quite difficult, because there are billions of combinations. The real number of combinations is $4.3252*10^{19}$ \cite{korf01} and all of them can be reached from any starting combination. The puzzle is successfully resolved when a sequence of moves is applied such that all subcubes are matched to each other by their color on each side of the cube. By estimation in \cite{korf01}, resolutions sequences variate from 50 to 100 moves when the cube is well scrambled.

@@

The rest of this paper is organized as follows: Section 2 describes theoretical details. Section 3 presents the proposed modifications into a practical software example. Section 4 is devoted to some experiments and results. Finally, Section 5 concludes and presents some ideas for further research.

@@

Genetic algorithms are the base of the genetic programming which is used in this research. Each element of the solution vector actually is an operation applied over the state of the Rubik's cube. Ordered set of such instructions is actually an algorithmic program. Because there is no direct intermediate relation between the individuals in a particular population genetic algorithms are perfect for implementation in parallel computing or distributed computing. Population of the genetic algorithm can be easily divided is many sub-population and it can be distributed on many processors/cores or even heterogeneous computers in a cluster. Separation of the global population is the preferred approach, but in cases where only the fitness value calculation is time consuming population is kept in the central processor/computer and only fitness value calculation is sent to the other contributing processors/computers.

@@

Single cut point is selected as crossover population, but other options \cite{poli01} are also applicable. As mutation operator random change of a single instruction is selected. Selection is done by randomly selected parents, but elitism rule is applied. For the evaluation of the newly created individuals instructions encoded in the individual are applied over the scrambled cube. After that the state of the cube is compared with the target state (cube in the solved state). The listing in Fig.\ref{fig01} shows the proposed in this paper modification of fitness evaluation function. For each pairs of cube's sides Euclidean distance is calculated. After that according Hausdorff distance rules the maximum of the minimums is found. Evaluated fitness value is positive, because the Euclidean distance is calculated with positive integers (cube's colors are mapped to integers) and the Hausdorff distance is calculation of a maximum of the minimums. The puzzle is as better solved as the fitness value is smaller.

@@

Fig.\ref{fig03} shows that when star topology is used the advantage of Hausdorff modification is not so big, but as it is shown in Fig.\ref{fig04} when incident nodes participation is used the proposed modification leads to convergence seep-up.

@@

The experiments show that the addition of Hausdorff distance component improves the performance of the genetic algorithm. Calculation of the Hausdorff distance is a little bit slower than the calculation of the Euclidean distance, but better solution fitness estimation generally leads to genetic algorithm convergence improvement. Investigation in the field of the soft computing can be done for the proposed fitness value evaluation as it was done in \cite{angelova01}.

As further research it will be interesting evaluation of the fitness value to be additionally filtered with Kalman filter \cite{alexandrov01} for example. The other interesting direction would be imvolwment of the artificial neural networks such as \cite{tashev01,atanasova01} for preliminary solutions evaluation.

>>>>

A parallel implementation of a genetic algorithm-based solver of the Rubik's cube was implemented by Balabanov in \cite{balabanov01} and it was presented in \cite{balabanov02}. Rubik's cube was invented and introduced by Erno Rubik in the 70s of the 20th century. After its creation the cube became the most popular combinatorial puzzle all over the world. In its original version it has 3x3x3 cubical segments. There are stickers in six different colors on each subcube square of the exposed sides. Each of the six planes (3x3x1) can be rotated in 90, 180, 270 or 360 degrees, relative to the other part of the puzzle. In the original initial state all sides of the cube are in single color. Scumbling of the puzzle is done by many random rotations of the (3x3x1) sides. The optimization task aims to restore the cube in its original state. Such combinatorial optimization problem is quite difficult because there are billions of combinations. The real number of combinations is $4.3252*10^{19}$ \cite{korf01} and all of them can be reached from any starting combination. The puzzle is successfully resolved when a sequence of moves is applied such that all subcubes are matched to each other by their color on each side of the cube. According to an estimation in \cite{korf01} resolutions sequences variate from 50 to 100 moves when the cube is well scrambled.

@@

The rest of this paper is organized as follows: Section 2 briefly describes theoretical details. Hence, section 3 presents the proposed modifications into a practical software example and section 4 is devoted to some experiments and results. Finally, Section 5 concludes and presents some ideas for further research.

@@

Genetic algorithms are the base of the genetic programming which is used in this research. Each element of the solution vector actually is an operation applied over the state of the Rubik's cube. Ordered set of such instructions is actually an algorithmic program. Because there is no direct intermediate relation between the individuals in a particular population genetic algorithms are highly appropriate for implementation in parallel computing or distributed computing. Population of the genetic algorithm can be easily divided is many sub-population and it can be distributed on many processors/cores or even heterogeneous computers in a cluster. Separation of the global population is the preferred approach, but in cases where only the fitness value calculation is time consuming, population is kept in the central processor/computer and only fitness value calculation is sent to the other contributing processors/computers.

@@

Single cut point is selected as crossover population, but other options \cite{poli01} are also applicable. As mutation operator random change of a single instruction is selected. Selection is done by randomly selected parents, but elitism rule is applied. For the evaluation of the newly created individuals instructions encoded in the individual are applied over the scrambled cube. After that the state of the cube is compared with the target state (cube in the solved state). The listing in Fig.\ref{fig01} shows the proposed in this paper modification of fitness evaluation function. For each pairs of cube's sides Euclidean distance is calculated. After that according to Hausdorff distance rules the maximum of the minimums is found. Evaluated fitness value is positive because the Euclidean distance is calculated with positive integers (cube's colors are mapped to integers) and the Hausdorff distance is a calculation of a maximum of the minimums. The puzzle is as better solved as the fitness value is smaller.

@@

Fig.\ref{fig03} shows that when star topology is used the advantage of Hausdorff modification is not so great, but as it is shown in Fig.\ref{fig04} when an incident nodes participation is used the proposed modification leads to convergence seep-up.

@@

The experiments show that the addition of Hausdorff distance component improves the performance of the genetic algorithm. A calculation of the Hausdorff distance is a little bit slower than the calculation of the Euclidean distance, but better solution fitness estimation generally leads to genetic algorithm convergence improvement. Further investigations in the field of the soft computing can be done for the proposed fitness value evaluation as it was done in \cite{angelova01}.

As further research it will be interesting an evaluation of the fitness value to be additionally filtered with Kalman filter \cite{alexandrov01}, for example. Another interesting direction would be an involvement of the artificial neural networks such as \cite{tashev01,atanasova01} for preliminary solutions evaluation.
